---
title: "Instructions for use"
author: "Drew Tyre"
date: "June 1, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pRoot <- rprojroot::find_rstudio_root_file
library(dplyr)
library(purrr)
library(tidyr)

```

First load the .RData object. I have it stashed in a subdirectory "data".

```{r}
load(pRoot("data","COMPADRE_v.3.2.1.RData"))
```

The first thing I want to do is find out how many matrices there are for a particular species. 

```{r}

compadre$metadata %>% filter(Genus == "Astragalus") %>% group_by(MatrixStartYear) %>% summarize(nMatrices = n()) %>% print(n = 1000)
```

So we have matrices for *Astrgalus sp.* from 1987 through 2008 already in the database. I'm going to try that again, but this time zoon in to 1991 and ask for the MatrixPopulation names. 

```{r}
library(dplyr)
compadre$metadata %>% filter(Genus == "Astragalus", MatrixStartYear == 1991) %>% select(MatrixPopulation, StudiedSex, MatrixComposite )
```

That's cool there are a bunch of sites in there that are not part of the long term study. Now I want to pull out the individual matrices for Haynes Creek. There's a script called "subsettingCOMPADRE.R" in the scripts directory from the github site, so I'll look at that. Seems pretty easy, the idea is to use the rownames from the return value of `subset()` to subset the mat object. 

```{r}
tempMetadata <- subset(compadre$metadata,MatrixComposite == "Individual" & 
    MatrixPopulation == "Haynes Creek" & Genus == "Astragalus")

# Now I can use the row names from the subsetted dataframe to subset the matrices.

keep <- as.numeric(rownames(tempMetadata))
tempMat <- compadre$mat[keep]
length(tempMat)

tempMat[[1]]
```

That's cool. But doing it this way means the result looses all information about where it came from! Now I want to know what the classes are called. In the 2015 paper they mention a variable `MatrixClassAuthor` has this information.

```{r}
names(tempMetadata)
select(tempMetadata,MatrixCriteriaOntogeny,MatrixCriteriaSize,MatrixCriteriaAge)
select(tempMetadata,MatrixDimension, MatrixFec, MatrixClassAuthor)

```

There is no such beast. But! there is another component of the compadre object called matrixClass

```{r}
length(compadre$matrixClass)
nrow(compadre$metadata)
astra1991 <- compadre$metadata %>% filter(Genus == "Astragalus", MatrixStartYear == 1991) %>% select(MatrixPopulation, StudiedSex, MatrixComposite )
row.names(astra1991)
tempClasses <- compadre$matrixClass[keep]
tempClasses[[1]]
tempClasses[[16]]
```

The next question I'm wondering about is the proportion of all plant matrices that have a dormant class. That is, one or more classes that equal `dorm` in `MatrixClassOrganized`.

```{r}
dorm <- function(x){
  # x is a list with a data.frame having three columns
  # I want to return true if one or more rows of MatrixClassOrganized is "dorm"
  sum(x$MatrixClassOrganized == "dorm") > 0
}
tempMetadata <- compadre$metadata
tempMetadata %>% mutate(hasDorm= map_lgl(compadre$matrixClass,dorm)) %>% count(GrowthType, hasDorm) %>% 
    spread(hasDorm, n, fill = 0)
```

So quite a few herbaceous perennial matrices have dormancy. But how many species? Is there a field that uniquely identifies a species? 

```{r}
tbl_df(compadre$metadata) %>% mutate(hasDorm= map_lgl(compadre$matrixClass, dorm)) %>% 
  group_by(SpeciesAccepted) %>%
  summarize(GrowthType = first(GrowthType),
            hasDorm = sum(hasDorm) > 0) %>%
  count(GrowthType, hasDorm) %>% 
  spread(hasDorm, n, fill = 0)
```

  So that's more interesting! about 10% of herbaceous perennial species have evidence of dormancy. The proportion of species with dormancy is much lower in other groups. 
